### 2020-08-15 code/supplemental_v2.Rmd

s4plot <- function(x) {
  a <-  x[,"start1"]
  b <- x[,"end1"]
  c <- x[,"start2"]
  d <- x[,"end2"]
  e <- x[,"start3"]
  f <- x[,"end3"]
par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in pathorder[x[,"start1"]:x[,"end1"]]) {
  pathname <- as.data.frame(strsplit(n, " ")) ## split string into word
  pathname <- pathname[,1] ## get the character vectors
  with(selected_path %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
path1 <- paste0(pathname[1], " ", pathname[2], " ", pathname[3], " ", pathname[4], " ")
path1 <- gsub("NA", "", path1)
path2 <- paste0(pathname[5], " ", pathname[6], " ", pathname[7], " ", pathname[8], " ")
path2 <- gsub("NA", "", path2)
legend("top", inset = c(0.01), legend=path1, bty = "n", horiz = T, cex = 0.9)
legend("top", inset = c(0.06), legend=path2, bty = "n", horiz = T, cex = 0.9)
}

@@

a = 1
b = 6
c = 7
d = 12
e = 13
f = 15

for(p in 1:nrow(startingpointlist)-1) {
  data <- startingpointlist[p,]
  s4plot(x = data)
}

# par(mfrow = c(2,3), mai = c(0.1, 0.1, 2, 0.1))
# par(mar = c(2,2,0,0))
# for(n in genes[7:12]) {
# with(selected_mut %>% 
#        filter(variable == n),
#      scatter3D(plot_y, -plot_z, plot_x, 
#                                 bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
#                                 theta = 0, phi = 65, scale = F,
#                xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
#                                 xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
#                                 colvar = NULL))
# title(xlab="UMAP 2", line=0)
# legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
# }
# par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
# par(mar = c(2,2,0,0))
# for(n in genes[13:15]) {
# with(selected_mut %>% 
#        filter(variable == n),
#      scatter3D(plot_y, -plot_z, plot_x, 
#                                 bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
#                                 theta = 0, phi = 65, scale = F,
#                xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
#                                 xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
#                                 colvar = NULL))
# title(xlab="UMAP 2", line=0)
# legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
# }

```
**Supplementary Figure S4. Projection of pathway disruptions onto the UMAP.** Each panel shows one of 377 Reactome
pathways projected on the UMAP from **Figure 1D of the main text** ordered by the hierarchical pathway order 
(column) from **Figure 2C of the main text**.

```{r supplementary-figure-s4, message = FALSE, echo = FALSE, out.width = '90%', fig.align= 'center', eval = FALSE}
load("../data/fig_s4_ggplot.Rda")

ggplot_umap_1
ggplot_umap_2
ggplot_umap_3
ggplot_umap_4
ggplot_umap_5
ggplot_umap_6
ggplot_umap_7
ggplot_umap_8
ggplot_umap_9
ggplot_umap_10
ggplot_umap_11
ggplot_umap_12
ggplot_umap_13
ggplot_umap_14
ggplot_umap_15
ggplot_umap_16

# ggplot(selected_path, aes(x = x_2d, y = y_2d, colour = as.character(value))) + 
#   geom_point(data = umap_bg, aes(x = x_2d, y = y_2d), colour = "grey", size = 0.5) +
#   geom_point(size = 0.5) +
#   theme_classic() +
#   coord_fixed() +
#   xlab("UMAP 1") + ylab("UMAP 2") +
#   scale_colour_manual(values = rep("firebrick3", 20)) +
#   theme(axis.text.x = element_blank(),
#     axis.text.y = element_blank(),
#     axis.title.x = element_blank(),
#     axis.title.y = element_blank(),
#     axis.ticks = element_blank(),
#     axis.line = element_blank(),
#     legend.position = "none") +
#   facet_wrap(~ factor(variable, levels = toppath),
#                       ncol = 5) ## facet by pathways

```

@@

## class 7 and 8 distinguishing genes
genes <- c("PIK3CA", "KRAS", "TP53")

@@

```{r supplemental-s5-plot, fig.width=9}
umap_cancertype <- data.table::fread("../data/umap_3d_coors.tsv")

umap_cancertype$clust_knn <- as.character(umap_cancertype$clust_knn)

clusters178 <- umap_cancertype %>% 
       mutate(color = case_when(clust_knn %in% c("1", "7", "8") ~ clust_knn,
                                  TRUE ~ "Other"))

I_outlines <- c("black", "black", "black", "#00000000")
names(I_outlines) <- c("1", "7", "8", "Other")

I_colors <- c("gold2", "gray51", "springgreen4", "#00000011")
names(I_colors) <- c("1", "7", "8", "Other")

load("../data/s5_data.Rda")
genes <- as.character(unique(s5_data$variable))

par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
plot.new()
with(clusters178, 
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = I_colors[color], col = I_outlines[color],
                                pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
legend("top", inset = c(0.05), legend=names(I_colors[c(1:3)]), col=I_colors[c(1:3)], pch = 20, bty = "n", horiz = T, cex = 1.2)
title(xlab="UMAP 2", line=0)
plot.new()

for(n in genes) {
with(s5_data %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}
```
**Supplementary Figure S5. Distribution of oncogenes on UMAP clusters.** The left panel shows the identity of 
clusters 1, 7 and 8. The multipanel on the right shows one of the three oncogenes projected onto the tumors of 
the UMAP from **Figure 2C of the main text**.

>>>>

s4plot <- function(x) {
  a <-  x[,"start1"]
  b <- x[,"end1"]
  c <- x[,"start2"]
  d <- x[,"end2"]
  e <- x[,"start3"]
  f <- x[,"end3"]
  path01<- pathorder[a:b]
  path02<- pathorder[c:d]
  path03<- pathorder[e:f]
par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in path01) { #pathorder[x[,"start1"]:x[,"end1"]]
  pathname <- as.data.frame(strsplit(n, " ")) ## split string into word
  pathname <- pathname[,1] ## get the character vectors
  with(selected_path %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
path1 <- paste0(pathname[1], " ", pathname[2], " ", pathname[3], " ", pathname[4], " ")
path1 <- gsub("NA", "", path1)
path2 <- paste0(pathname[5], " ", pathname[6], " ", pathname[7], " ", pathname[8], " ")
path2 <- gsub("NA", "", path2)
legend("top", inset = c(0.01), legend=path1, bty = "n", horiz = T, cex = 0.9)
legend("top", inset = c(0.06), legend=path2, bty = "n", horiz = T, cex = 0.9)
}

@@

data <- startingpointlist[1,]
s4plot(x = data)
data <- startingpointlist[2,]
s4plot(x = data)
data <- startingpointlist[3,]
s4plot(x = data)
data <- startingpointlist[4,]
s4plot(x = data)
data <- startingpointlist[5,]
s4plot(x = data)
data <- startingpointlist[6,]
s4plot(x = data)
data <- startingpointlist[7,]
s4plot(x = data)
data <- startingpointlist[8,]
s4plot(x = data)
data <- startingpointlist[9,]
s4plot(x = data)
data <- startingpointlist[10,]
s4plot(x = data)
data <- startingpointlist[11,]
s4plot(x = data)
data <- startingpointlist[12,]
s4plot(x = data)
data <- startingpointlist[13,]
s4plot(x = data)
data <- startingpointlist[14,]
s4plot(x = data)
data <- startingpointlist[15,]
s4plot(x = data)
data <- startingpointlist[16,]
s4plot(x = data)
data <- startingpointlist[17,]
s4plot(x = data)
data <- startingpointlist[18,]
s4plot(x = data)
data <- startingpointlist[19,]
s4plot(x = data)
data <- startingpointlist[20,]
s4plot(x = data)
data <- startingpointlist[21,]
s4plot(x = data)
data <- startingpointlist[22,]
s4plot(x = data)
data <- startingpointlist[23,]
s4plot(x = data)
data <- startingpointlist[24,]
s4plot(x = data)
data <- startingpointlist[25,]
s4plot(x = data)

par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in pathorder[376:377]) { 
  pathname <- as.data.frame(strsplit(n, " ")) ## split string into word
  pathname <- pathname[,1] ## get the character vectors
  with(selected_path %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
path1 <- paste0(pathname[1], " ", pathname[2], " ", pathname[3], " ", pathname[4], " ")
path1 <- gsub("NA", "", path1)
path2 <- paste0(pathname[5], " ", pathname[6], " ", pathname[7], " ", pathname[8], " ")
path2 <- gsub("NA", "", path2)
legend("top", inset = c(0.01), legend=path1, bty = "n", horiz = T, cex = 0.9)
legend("top", inset = c(0.06), legend=path2, bty = "n", horiz = T, cex = 0.9)
}
```
**Supplementary Figure S4. Projection of pathway disruptions onto the UMAP.** Each panel shows one of 377 Reactome
pathways projected on the UMAP from **Figure 1D of the main text** ordered by the hierarchical pathway order 
(column) from **Figure 2C of the main text**.

@@

## class 7 and 8 distinguishing genes
genes <- c("AKT2","BCL2","CTNNB1","ERBB2","KRAS","PIK3CA","CDKN2C","PTEN","RB1","SMARCA4","TP53","VHL")

@@

```{r supplemental-s5-plot, fig.width=9}
# umap_cancertype <- data.table::fread("../data/umap_3d_coors.tsv")
# 
# umap_cancertype$clust_knn <- as.character(umap_cancertype$clust_knn)
# 
# clusters178 <- umap_cancertype %>% 
#        mutate(color = case_when(clust_knn %in% c("1", "7", "8") ~ clust_knn,
#                                   TRUE ~ "Other"))
# 
# I_outlines <- c("black", "black", "black", "#00000000")
# names(I_outlines) <- c("1", "7", "8", "Other")
# 
# I_colors <- c("gold2", "gray51", "springgreen4", "#00000011")
# names(I_colors) <- c("1", "7", "8", "Other")
# 
# load("../data/s5_data.Rda")
# genes <- as.character(unique(s5_data$variable))
# 
# par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
# par(mar = c(2,2,0,0))
# plot.new()
# with(clusters178, 
#      scatter3D(plot_y, -plot_z, plot_x, 
#                                 bg = I_colors[color], col = I_outlines[color],
#                                 pch = 21, cex = 0.8, lwd = 0.2,
#                                 theta = 0, phi = 65, scale = F,
#                                 xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
#                                 colvar = NULL))
# legend("top", inset = c(0.05), legend=names(I_colors[c(1:3)]), col=I_colors[c(1:3)], pch = 20, bty = "n", horiz = T, cex = 1.2)
# title(xlab="UMAP 2", line=0)
# plot.new()

load("../data/s5_data.Rda")
genes <- c("AKT2","BCL2","CTNNB1","ERBB2","KRAS","PIK3CA","CDKN2C","PTEN","RB1","SMARCA4","TP53","VHL")

par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in genes[1:6]) {
with(s5_data %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}

par(mfrow = c(2,3), mai = c(0.1, 0.1, 0.1, 0.1))
par(mar = c(2,2,0,0))
for(n in genes[7:12]) {
with(s5_data %>% 
       filter(variable == n),
     scatter3D(plot_y, -plot_z, plot_x, 
                                bg = "grey", pch = 21, cex = 0.8, lwd = 0.2,
                                theta = 0, phi = 65, scale = F,
               xlim = c(-6, 4.5), zlim = c(-4.5, 4), ylim = c(-4, 5),
                                xlab = "", ylab = "UMAP 3", zlab = "UMAP 1",
                                colvar = NULL))
title(xlab="UMAP 2", line=0)
legend("top", inset = c(0), legend=n, bty = "n", horiz = T, cex = 1.2)
}

```
**Supplementary Figure S5. Distribution of oncogenes/tumor suppressors on UMAP clusters.** Each panel shows shows one of the twelve oncogenes/tumor suppressors projected onto the tumors of the UMAP from **Figure 2C of the main text**.

### 2020-08-15 code/survival/survfig.Rmd

#### Figure 8
```{r, message = FALSE, fig.width = 9}
mylayout = '
ABB
ABB
#BB
'

((pcanc+mytheme) + (pclass+mytheme))+plot_layout(design=mylayout) +plot_annotation(tag_levels = "A")
```

>>>>

#### Figure 8
```{r, message = FALSE, fig.width = 3}
pcanc+mytheme
```

#### Figure 9
```{r, message = FALSE, fig.width = 9, fig.height= 9}
kelly.colours <- c("gray95", "gray13", "gold2", "plum4", "darkorange1",
                   "lightskyblue2", "firebrick", "burlywood3", "gray51", "springgreen4",
                   "lightpink2", "deepskyblue4", "lightsalmon2", "mediumpurple4",
                   "orange", "maroon", "yellow3", "brown4", "yellow4", "sienna4", "chocolate", "gray19")

cluster.colours <- c("grey", kelly.colours[c(3:12)])
names(cluster.colours) <- c(0:10)

kclassplot = function(dfsamps,tisnames){
	kclnames = grep("k\\[",names(dfsamps), value=TRUE)
	dfclas = dfsamps[,kclnames]
	dfclas = stack(dfclas)
	names(dfclas) = c("value","coefname")
	dfclas$class = as.factor(as.numeric(gsub("k\\[(.*),(.*)\\]","\\2",dfclas$coefname)))
	dfclas$tiss = tisnames[as.numeric(gsub("k\\[(.*),(.*)\\]","\\1",dfclas$coefname))]
	ggplot(dfclas,aes_(quote(value),quote(class), fill = quote(class))) + geom_density_ridges() +
		facet_wrap(. ~ tiss,nrow=3) + theme_minimal() + scale_fill_brewer(palette = "Paired") +
		xlim(0,3) + theme(legend.position = "none") + ggtitle("Class Specific Coefficients")
}

pclass = kclassplot(dfsamps,levels(cancerdat$type))

pclass+mytheme+scale_fill_manual(values = cluster.colours)
```
